---
title: "R Notebook"
output: html_notebook
---

## Project Info - Assignment 3
Nadejda Boev (20056079)
Due Date - 2022/01/26
Github - https://github.com/16nbb1/Biol432_A3

### Data Input and libraries
Loading in libraries I'll Needd

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(stringr)

```

Loading in dataset

```{r}
(df = read.csv("./InputData/FallopiaData.csv"))
```

### Data Wrangling

1. Removing rows with ‘Total’ biomass < 60
* Filtering for rows that have a biomass of 60 or less
2. Make a new column TotalG, which converts the ‘Total’ column from mg to grams AND replace Total with TotalG, and add it to the dataset.
* Removing Total using NULL 
3. Reorder the columns so that they are in the order: ‘TotalG’, ‘Taxon’, ‘Senario’, ‘Nutrients’ AND Remove the other columns

NEEED TO DOUBLE CHECK MG VS G

```{r}
df_filt = df %>% 
  filter(Total <= 60) %>% 
  mutate(TotalG = Total / 1000, Total = NULL) %>% 
  select(TotalG, Taxon, Scenario, Nutrients) 
```

### Function: Basic Stats

Input: 
* A vector of data to process (e.g. column from a data.frame object) 
* A string that defines what calculation to perform.
  * Options: "Average", "Sum", "Observations"
  * All other inputs recieve the following message: "ERROR input"

Output: 
* 

o	Write some R code that uses your function to count the total number of observations in the 'Taxon' column.
o	Write some R code that uses your function  to calculate the average TotalG for each of the two Nutrient concentrations.


```{r}
basics <- function (in_comm, in_vec, datafram) {

  if(in_comm == 'Average') {
    print(paste0('Average: ', mean(in_vec, na.rm=T)))
  } else if (in_comm == 'Sum') {
    print(paste0('Sum: ', sum(in_vec, na.rm=T)))
  } else if (in_comm == 'Observations') {
    print(paste0('Number of observations: ',length(in_vec)))
  } else {
    print('INPUT ERROR')
  }
  
  if(any(str_detect(names(datafram), "Taxon")) == TRUE) {
    print(paste0('Number of observations in Taxon column: ',length(datafram$Taxon)))
  } else {
    print('ERROR : No taxon column')
  }
    
  if(any(str_detect(names(datafram), "TotalG")) == TRUE) {
      groupby_nut = datafram %>% 
      group_by(Nutrients) %>% 
      summarize(Mean = mean(TotalG))
      
      print(paste0('Average biomass for high nutrients:  ', signif(groupby_nut[groupby_nut$Nutrients == 'high',]$Mean, digits = 3), 'g'))
      print(paste0('Average biomass for low nutrients:  ', signif(groupby_nut[groupby_nut$Nutrients == 'low',]$Mean, digits = 3), 'g'))
 
     } else {
    print('ERROR : No TotalG column')
  }


  
}

```

Testing inputs

Successful cases : 

```{r}
basics('Average', df_filt$TotalG,  df_filt)
basics('Sum', df_filt$TotalG,  df_filt)
basics('Observations', df_filt$TotalG,  df_filt)

```

Other cases : 

```{r}
# Spelling error
basics('Observaions', df_filt$TotalG,  df_filt)

# Throwing error when wrong dataaframe is used (df does not have a TotalG column)
basics('Observations', df_filt$TotalG,  df)
```

Saving the "clean" dataset

```{r}
write.csv(df_filt,"./Output/FallopiaData_60less.csv")
```



